Supplemental Material for submission ``Do Objects in Real-World Images Have a Canonical Name?''

### 2019-05-17 papers/EMNLP2019/analysis.tex
looked at freq. thresholds, separated type counts (histograms) and agreement figures in analysis

In this section, we investigate to what extent names annotated in VisualGenome and elicited in ManyNames can be considered canonical, i.e. to what extent speakers agree in their naming choices.
Whereas traditional picture naming studies typically use a prototypical image per category (see Figure~\ref{fig:picture_naming}) and, hence, are mostly interested in the agreement on concept or category-level, we carry out an analysis on two different levels: First, we will look at instances and see to what extent names overlap for the same object. 
Second, we will use the existing annotation of names in VisualGenome to analyze agreement on the level of categories.
\begin{figure}[t]
	\centering
	\includegraphics[scale=.5]{figures/snodgrass_vanderwart_cake_042.png}
	\caption{Example of a picture of \textsl{cake} used in traditional picture naming studies (REF to Vanderwart \& Snodgrass) \label{fig:picture_naming}}
\end{figure}

@@

\begin{itemize}
\item \textbf{\% top}: the average relative frequency of the most frequent response (shown in percent)
\item \textbf{$H$}: the $H$ agreement measure used previously in the psycholinguistic literature
\cs{How is this defined?}
\item \textbf{N}: the average number of types in the response set of ManyNames
\item \textbf{N$_{>1}$}: the average number of types, excluding types that have been annotated only once
\cs{alternatively we could show a plot going from 1 to, let's say, $>10$}
\item \textbf{top=VG}: the proportion of items where the top response in ManyNames corresponds to the VisualGenome name
\item \textbf{\% VG}: the average relative frequency of the VisualGenome name in the response set

@@

\begin{table*}
\footnotesize
\begin{tabular}{p{1.3cm}cccccc|cccccc}
\toprule
 & \multicolumn{6}{c|}{Instance-level agreement} & \multicolumn{6}{c}{Class-level agreement}\\ 
           domain & \% top &    $H$ &    N & N$_{>1}$ & top=VG &  \% VG & \% top &    $H$ &     N & N$_{>1}$ & top=VG &  \% VG \\
\midrule
            all &  69.7 (22.2) &  1.3 &  5.7 &   2.9 &  72.8  &  58.7  &   52.6 (21.1) &  2.4 &   62.2 &  29.7 &  32.7 (46.9) &  23.5 (27.0)  \\
 \midrule
         people &  51.9 (18.1) &  2.1 &  8.6 &   4.3 &         49.8 &         32.3 &     43.1 (14.9) &  2.9 &  104.4 &  53.4 &         24.4 &         13.2  \\
 animals\_plants &  91.3 (12.9) &  0.4 &  2.7 &   1.5 &         93.8 &         88.0 &     67.9 (23.5) &  1.5 &   26.7 &  12.4 &         29.5 &         26.2  \\
       clothing &  63.9 (17.9) &  1.6 &  6.4 &   3.2 &         70.2 &         52.6 &     49.3 (16.6) &  2.6 &   71.7 &  34.1 &         40.5 &         26.1  \\
       vehicles &  72.0 (19.5) &  1.1 &  4.7 &   2.4 &         71.1 &         60.2 &    54.4 (17.4) &  2.1 &   70.4 &  33.2 &         21.4 &         21.3  \\
      buildings &  66.9 (20.5) &  1.5 &  6.9 &   3.0 &         72.6 &         55.5 &      46.7 (18.2) &  3.0 &   68.5 &  31.3 &         32.3 &         22.3  \\
           home &  66.4 (20.5) &  1.5 &  6.3 &   3.1 &         78.5 &         58.8 &     49.6 (18.7) &  2.8 &  103.2 &  48.8 &         45.9 &         30.1  \\
           food &  71.3 (21.1) &  1.3 &  5.5 &   2.9 &         62.9 &         52.1 &     47.3 (19.5) &  2.5 &   32.1 &  15.2 &         31.1 &         20.8  \\
\bottomrule
\end{tabular}\caption{Agreement in naming measured on the level of instances and on the level of VG classes (i.e.\ after grouping objects by their VG synset) \cs{\textbf{std=?})}}
\label{tab:agree}
\end{table*}

>>>>

%\gbt{Points to discuss in the paper, as discussed with Carina May 16:}
%
%\begin{itemize}
%\item Super high variance in the agreement. Higher agreement for instances, lower for classes. Suggests systematicities at the instance level that do not depend on the class. Hypothesis, supported by qualitative analysis (Figure 3): relevance of visual perceptual factors. E.g.: saliency, background/foreground (desk-keyboard; bridge-train); ``angle'' from which the object is shown (see girl-t-shirt example); \dots In addition, we also find aspects discussed by Psycholing, but at the level of the instance: Typicality (see truck-bus example, bridge-dock, bench-seat). 
%\item Forget about WordNet.
%\item Implications for lang\&vision: 1) synset classification won't do (if the goal is to predict/label an object); 2) name classification won't do either; 3) ``mistakes'' done by models are also done by humans -- role of referential uncertainty.
%\end{itemize}

In this section, we investigate to what extent names annotated in VisualGenome and elicited in ManyNames can be considered canonical, i.e. to what extent speakers agree in their naming choices.
Whereas traditional picture naming studies typically use a prototypical image per category (see Figure~\ref{fig:picture_naming}) and, hence, are mostly interested in the agreement on concept or category-level, we carry out an analysis on two different levels: First, we will look at instances and see to what extent names overlap for the same object. 
Second, we will use the existing annotation of names in VisualGenome to analyze agreement on the level of classes.
\begin{figure}[t]
	\centering
	\includegraphics[scale=.5]{figures/snodgrass_vanderwart_cake_042.png}
	\caption{Example of a picture of \textsl{cake} used in traditional picture naming studies (REF to Vanderwart \& Snodgrass) \label{fig:picture_naming}}
\end{figure}

\subsection{Names, instances, classes}

First of all, to see whether objects in ManyNames bear a canonical name, we simply count how many different names (i.e.\ types) are given to instances, and to all instances that have the same synset annotated in VisualGenome. 
As the data is collected via crowdsourcing and a certain amount of noise is to be expected, we apply different frequency thresholds on the response sets for instances and synsets.
Figure \ref{fig:ntypes} shows the cumulative histograms for type counts, obtained with different thresholds (from 1 to 6).
Without any frequency thresholding, the proportion of instances and classes that have a single name annotated is small, i.e.\ below 10\% in both cases. When raising the threshold up to a minimum of 4 occurences in the response set  for instances (meaning more than 10\% of the 36 annotations), the proportion of objects that really only have a single name annotated is considerably higher, but still below 50\%.
Thus, as expected in a free annotation scenario, the MN data contains a certain number of low-frequency responses.
The fact that even after applying a relatively strict frequency filter most objects have more than one name annotated indicates that there is a non-negligible  level of variation when different speakers name the same object.

The amount of variation increases substantially when looking at the number of names we find for different synsets in Visual Genome, as is indicated by the histogram for type counts of synsets in Figure \ref{fig:ntypes}.
Here, we observe that the rise of the curve is much less steep than for instances and, generally, there are many synsets with more than 10 or even 20 different name types, even after applying the frequency threshols.

\begin{figure*}
\begin{minipage}[b]{0.4\linewidth}
\includegraphics[scale=.4]{Figures/types_instances.png}
\end{minipage}
\begin{minipage}[b]{0.6\linewidth}
\includegraphics[scale=.4]{Figures/types_classes.png}
\end{minipage}
 \caption{\label{fig:ntypes} Cumulative histograms for number of types found for instances and classes, based on different frequency tresholds (applied on the level of instances and classes respectively)}
\end{figure*}

@@

\begin{itemize}
\item \textbf{\% top}: the average relative frequency of the most frequent response (shown in percent)
\item \textbf{$H$}: the $H$ agreement measure used previously in the psycholinguistic literature
\cs{How is this defined?}
%\item \textbf{N}: the average number of types in the response set of ManyNames
%\item \textbf{N$_{>1}$}: the average number of types, excluding types that have been annotated only once
\cs{alternatively we could show a plot going from 1 to, let's say, $>10$}
\item \textbf{top=VG}: the proportion of items where the top response in ManyNames corresponds to the VisualGenome name
\item \textbf{\% VG}: the average relative frequency of the VisualGenome name in the response set

@@

\begin{table*}
\footnotesize
%\begin{tabular}{p{1.3cm}cccccc|cccccc}
%\toprule
% & \multicolumn{6}{c|}{Instance-level agreement} & \multicolumn{6}{c}{Class-level agreement}\\ 
%           domain & \% top &    $H$ &    N & N$_{>1}$ & top=VG &  \% VG & \% top &    $H$ &     N & N$_{>1}$ & top=VG &  \% VG \\
%\midrule
%            all &  69.7 (22.2) &  1.3 &  5.7 &   2.9 &  72.8  &  58.7  &   52.6 (21.1) &  2.4 &   62.2 &  29.7 &  32.7 (46.9) &  23.5 (27.0)  \\
% \midrule
%         people &  51.9 (18.1) &  2.1 &  8.6 &   4.3 &         49.8 &         32.3 &     43.1 (14.9) &  2.9 &  104.4 &  53.4 &         24.4 &         13.2  \\
% animals\_plants &  91.3 (12.9) &  0.4 &  2.7 &   1.5 &         93.8 &         88.0 &     67.9 (23.5) &  1.5 &   26.7 &  12.4 &         29.5 &         26.2  \\
%       clothing &  63.9 (17.9) &  1.6 &  6.4 &   3.2 &         70.2 &         52.6 &     49.3 (16.6) &  2.6 &   71.7 &  34.1 &         40.5 &         26.1  \\
%       vehicles &  72.0 (19.5) &  1.1 &  4.7 &   2.4 &         71.1 &         60.2 &    54.4 (17.4) &  2.1 &   70.4 &  33.2 &         21.4 &         21.3  \\
%      buildings &  66.9 (20.5) &  1.5 &  6.9 &   3.0 &         72.6 &         55.5 &      46.7 (18.2) &  3.0 &   68.5 &  31.3 &         32.3 &         22.3  \\
%           home &  66.4 (20.5) &  1.5 &  6.3 &   3.1 &         78.5 &         58.8 &     49.6 (18.7) &  2.8 &  103.2 &  48.8 &         45.9 &         30.1  \\
%           food &  71.3 (21.1) &  1.3 &  5.5 &   2.9 &         62.9 &         52.1 &     47.3 (19.5) &  2.5 &   32.1 &  15.2 &         31.1 &         20.8  \\
%\bottomrule
%\end{tabular}

\begin{tabular}{lcccc|cccc}
\toprule
 & \multicolumn{4}{c|}{Instance-level agreement} & \multicolumn{4}{c}{Class-level agreement}\\ 
         domain &         \%top &    H &       top=VG &          \%VG &         \%top &    H &       top=VG &          \%VG \\
\midrule
            all &  78.5 (21.5) &  0.7 &  72.8 (44.5) &  65.1 (36.1) &  58.9 (21.9) &  1.8 &  32.7 (46.9) &  23.2 (27.1) \\
 animals\_plants &  95.8 (11.4) &  0.1 &         93.8 &         92.1 &  74.3 (22.5) &  1.0 &         29.5 &         25.8 \\
      buildings &  79.0 (21.1) &  0.7 &         72.6 &         64.8 &  53.5 (19.4) &  2.2 &         32.3 &         22.1 \\
           home &  76.3 (20.6) &  0.8 &         78.5 &         67.0 &  56.4 (20.3) &  2.1 &         45.9 &         29.9 \\
           food &  79.8 (20.5) &  0.7 &         62.9 &         57.0 &  52.7 (21.1) &  1.9 &         31.1 &         20.8 \\
       vehicles &  79.0 (19.9) &  0.6 &         71.1 &         65.6 &  58.6 (17.0) &  1.7 &         21.4 &         21.1 \\
         people &  64.4 (21.8) &  1.2 &         49.8 &         39.3 &  51.1 (19.6) &  2.2 &         24.4 &         12.9 \\
       clothing &  73.9 (18.5) &  0.9 &         70.2 &         60.2 &  56.2 (17.2) &  1.9 &         40.5 &         26.0 \\
\bottomrule
\end{tabular}

\caption{Agreement in naming measured on the level of instances and on the level of VG classes (i.e.\ after grouping objects by their VG synset), for filtered response sets (frequency threshold of 3)}
\label{tab:agree}
\end{table*}
